---
title: "How Can Political Parties Satisfy their Voters?"
author: "Timothy Regis"
date: 2020-09-24T21:13:14-05:00
bibliography: "references1.bib"
categories: ["R"]
tags: ["R Markdown", "cesR", "Voter Satisfaction"]
---

# Author(s)

Timothy Jordan Regis

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

```{r, warning = FALSE, message=FALSE, echo=FALSE}
library(tidyverse)
library(visdat)
library(skimr)
library(cesR)
library(labelled)
library(ggplot2)
```


# Abstract

This paper analyzes the most important issues to voters of the; Liberal, Conservative, and NDP political Parties in Canada with the intention of guiding politicians toward enacting better policies geared for voter satisfaction. From our study we found that Liberal voters could gain the most satisfaction from increased funding in the education and health care sectors, Conservative voters can benefit the most from solid increases in funding to the health care sector, and NDP voters can gain high amounts of satisfaction with increased funding in any or all of the chosen sectors, with health care appearing most prominent. This study is significant as voter satisfaction is extremely important in keeping elected Parties in office for future terms, thus politicians should try to pay very much attention to the things that their voters are most concerned with.

# Introduction

When a new Prime Minister is elected, their Party must review the promises they made while campaigning and decide which they will act upon and how they will fulfill them. These decisions play a key role in determining whether or not those who voted for the elected Party remain satisfied for their term. Since voter satisfaction is integral for reelection, the candidates and their respective Parties must pay a great deal of attention to the issues most important to their voters and focus their powers toward solving them if they hope to continue serving.

Using the data from the Canadian Election Study gathered by cesR, this paper highlights the issues at the forefront of voters' minds and how they want these issues to be addressed by the Party they voted for.
The data gathered was the ces2015_web study from the cesR package. After some filtering, the data was split into three groups according to each voter's response to the question "Who are you likely to vote for?" with the groups being, NDP, Liberal, and Conservative voters. Then, multiple grouped bar graphs were created to display the frequency of voters' responses to a variety of questions such as; "How much attention do you pay to education?", and "Should the federal government spend more, less, or about the same amount of money on education?" By studying the outcomes, we are able to gain an understanding of what issues voters of each Party consider the most important. From this, we can begin to make suggestions to each Party about which policies they should enact to result in the highest voter satisfaction possible. In doing so, this can help to secure a political Party member's position in office for multiple terms.

Our findings suggested to us that; Liberal voters can be highly satisfied by increased funding in the education and health care sectors, NDP voters are quite satisfied with increases in funding to all of our chosen sectors, with health care appearing to be the most satisfying, and lastly, Conservative voters can gain the most satisfaction from increased funding to the health care sector of the country. Still, however, further research into the data must be done to determine definitive answers to what policies each Party should focus on enacting.


# Data Identification


```{r, include=FALSE}
# devtools::install_github("hodgettsp/cesR")
```

```{r, fig.cap= 'Table 1', echo=FALSE}
get_cescodes()
```

From the list of available ces studies, we chose ces2015_web as the study for this paper, this was mainly due to the rich variety of questions answered in this survey. The results for this survey are displayed below.

```{r, echo=FALSE, include=FALSE}
get_ces('ces2015_web')
```

```{r, message=FALSE, echo=FALSE}
ces2015_web <- to_factor(ces2015_web)
head(ces2015_web)
```
The ces2015_web data displays rows representing each voter's list of responses to the questions in the survey, plus some data about how they operated the survey, including; browser type, start and finish times, and many more factors that likely won't be effective in this study, so it will require some initial cleaning.

Since by default, the ces data is not converted into a factor type, we first applied the function to_factor() from the labelled package to our ces2015_web data so that we can more easily work with it.
With a total column space of 580, the data is much too large to efficiently work with so, using the select function, we created a new variable, ces2015_web_wrk, which isolates the columns; province, attn_heal, attn_educ, attn_envi, vote_for, spnd_heal, spnd_educ, and spnd_envi
The corresponding questions to these columns are, in order:

Which province or territory are you currently living in? 
How much attention do you pay to health care? 
How much attention do you pay to education? 
How much attention do you pay to the environment? 
Which Party do you think you will vote for?
Should the federal government spend more, less, or about the same amount of money on health care?
Should the federal government spend more, less, or about the same amount of money on education?
Should the federal government spend more, less, or about the same amount of money on the environment?

I felt that these questions were of high importance on a nationwide level and that policies regarding them would be most likely to draw in voters due to this importance, thus these are the questions we will progress with for this study.

Examining Number of Voters for Each Party.
```{r, echo=FALSE}
ces2015_web_wrk <- ces2015_web %>%
  filter(vote_for %in% c('The Conservative Party', 'The Liberal Party', 'The New Democratic Party'))
table(ces2015_web_wrk$vote_for)
```

```{r, fig.cap='Graph 0.1', echo=FALSE}
ces2015_web_wrk %>% 
  ggplot(aes(x = vote_for)) + 
  geom_bar() + 
  labs(title = 'Number of Voters for Each Party', x = 'Political Party', y = 'Number of Voters') + 
  theme_minimal()
```
Next, we split the data up according to what Party each voter intends on voting for, using the most popular Parties, we are left with three groups of voters, Conservative Party voters, Liberal Party voters, and New Democratic Party voters. Since there is a difference in the number of voters in each group, we must keep this in mind as we progress, being sure not to make any strong comparisons between the groups before bringing them down to relative terms. Using this cleaned dataset, we are now able to begin examining the results of each voter's survey in our exploratory data analysis section up next.



# Exploratory Data Analysis

In this section, we will be analyzing the responses to our selected questions through multiple grouped bar graphs and attempting to determine which of these issues are most important to the voters of each Party and how they would like them to be addressed.

```{r, fig.cap='Graph 1.1', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = attn_educ, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Importance of Education to Voters', x = 'Attention to Education in the Country', y = 'Number of Voters') + 
  theme_minimal()
```

In our first graph, we examine responses to question attn_educ; 'How much attention do you pay to education?' As will be the case with the majority of these graphs, our data appear to be heavily grouped at the first three responses, a lot, a little, and none, which was to be expected.

Organized first by their response, and second by the political Party they intend to vote for, this grouped bar graph displays the quantity of each response recorded by the voters of each Party. The political Parties being voted for are identified by color, with red representing the Conservative Party, green representing the Liberal Party, and blue representing the New Democratic Party. 

As we can see, those who voted for the Liberal Party and the New Democratic Party responded saying they paid a lot of attention to education in the country in a higher proportion than those who vote for the Conservative Party as we notice more of them responding with only a little attention paid to education. This suggests to us that both the Liberal Party and the New Democratic Party should focus their power toward addressing the education system in Canada as this could seriously benefit voter satisfaction. On the other hand, the mixed results of Conservative voters indicate that policies regarding Canada's education system would not be most effective at spurring voter satisfaction.

```{r, fig.cap='Graph 1.2', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = spnd_educ, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Education Spending Recommendations by Voters', x = 'How should the Govt. Spend on Education ', y = 'Number of Voters') + 
  theme_minimal()
```
(Options Along X-Axis: 'Spend More', 'Spend Less', 'Spend About the Same as Now', 'Don't Know', 'Refuse to Answer', 'Seen But Not Answered', 'NA' )

When we plot the results to the question 'Should the federal government spend more, less, or about the same amount of money on education?' it becomes much more clear what type of policies each Party should enact. Due to the high proportion of Liberal and NDP voters that encourage more spending on the education system, it is highly recommended that, upon election, these Parties focus a portion of their spending on the education system to better satisfy the needs of their voters.
Policies such as; increasing the amount of money paid out by education plans like OSAP, building new schools, or renovating older schools have the ability to highly increase voter satisfaction, benefiting their chances of reelection.



```{r, fig.cap='Graph 2.1', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = attn_envi, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Importance of the Environment to Voters', x = 'Attention to the Environment in the Country', y = 'Number of Voters') + 
  theme_minimal()
```

Our next graph displays the results to the question attn_envi; 'How much attention do you pay to the environment?' in a similar fashion to the first graph wherein voters were organized by both their response to this question, as well as vote_for; "Which Party do you think you will vote for?"

The results suggest to us a couple of things regarding each Party; those voting for the Liberal Party are quite mixed in their attention to the environment in the Country, NDP voters more often expressed a lot of attention to the environment, and Conservative voters most often displayed a little attention to the environment. 

These results can help to guide politicians and policymakers of each Party when deciding on what policies to enact when elected. For example, given the turnout for Conservative Voters, the Conservative Party should not, upon election, announce a new and large carbon tax policy on non-electric vehicles, as this would not be effective at keeping voter satisfaction high.

```{r, fig.cap='Graph 2.2', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = spnd_envi, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Environmental Spending Recommendations by Voters', x = 'How should the Govt. Spend on the Environment', y = 'Number of Voters') + 
  theme_minimal()
```
In this graph we present the responses to the question spnd_envi; 'Should the federal government spend more, less, or about the same amount of money on the environment?' and given the responses to our question about the attention paid to the environment, this outcome was expected. 
As can be seen, both the Liberal Party and the New Democratic Party have the majority of their voters advocating for more spending in the environment sector whereas Conservative voters appear most often to suggest that roughly the same amount should be spent on the environment from past years. 
Thus, it is suggested that the Liberal and NDP Parties dedicate a larger portion of their spending toward benefiting the environment in Canada. For example, a project encouraging the use of renewable energy resources such as implementing a federal tax credit for people who purchase electric vehicles could be a powerful method of increasing voter satisfaction for the Party.


```{r, fig.cap='Graph 3.1', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = attn_heal, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Importance of Health Care to Voters', x = 'Attention to Health Care in the Country', y = 'Number of Voters') + 
  theme_minimal()
```
In this graph, results for the question attn_heal; 'How much attention do you pay to health care?' are displayed for each Party. As is apparent, health care is of extremely high importance to voters in each Party as all of them feature the most votes for a lot of attention paid to health care. This is not a surprising result at all however, as no matter how you vote, you will always need sufficient access to health care facilities to ensure that you can remain healthy throughout your life.
Based on this outcome, policymakers are suggested to address health care in the country no matter what Party they represent due to its high priority in the minds of voters. 


```{r, fig.cap='Graph 3.2', echo=FALSE}
ces2015_web_wrk %>%
  ggplot(aes(x = spnd_hlth, fill = vote_for)) + 
  geom_bar(color = 'black', position = 'dodge') + 
  labs(title = 'Health Care Spending Recommendations by Voters', x = 'How should the Govt. Spend on Education ', y = 'Number of Voters') + 
  theme_minimal()
```

The job of policymakers regarding the health care sector is clear when we analyze the results of the question spnd_hlth; 'Should the federal government spend more, less, or about the same amount of money on health care?' As it is quite obviously pointed out, all Parties should look to greatly increase their spending in the health care sector of Canada due to the exceptionally high demand for it. This could be accomplished in a number of ways such as; expanding overcrowded hospitals or designing new facilities geared toward aiding mental health. Policies like these would be extremely successful at increasing voter satisfaction across the country.

In situations like these where most voters have similar desires across Parties, if we were to see one Party promoting significantly higher increases to health care, it may be possible for them to steal voters that are on the fence from other Parties, possibly swaying the final results in their favor. Thus, it is encouraged that all Parties find some way to engage in the betterment of the health care sector in Canada.


# Conclusion & Discussion

From this analysis, we can begin to gain an understanding of what issues are most important to the Canadian voters of these three political Parties. It appears that for NDP voters, they pay great attention to all three categories; education, the environment, and health care while also advocating for increased funding in all of these sectors. Liberal Party voters expressed paying a lot of attention to both education and health care, while their attention to the environment appears mixed between a lot and a little attention paid. Despite this, Liberal voters indicated most often that they would like to see an increase in spending across all fields studied here. This can suggest that, while providing additional funding to the environment sector will likely increase voter satisfaction for the Liberal Party, it would not be as effective at doing so as providing funding to either the education or health care sectors in the country. Finally, Conservative Party voters seemed to be mixed between a lot and a little in their level of attention paid to both education and the environment. When it comes to funding, Conservative voters appeared mixed between spending more or about the same on education and showed more of a desire to keep funding for the environment at the level it is. It should be noted, however, when compared to the other voters' responses, taking into account the fact that there is a smaller pool of conservative voters, a relatively large proportion of Conservative voters suggested that funding for the environment be cut down. However, much like the other two Parties, voters for the Conservative Party expressed a lot of attention toward health care in Canada with the majority of them suggesting increased funding in this sector. By paying the most attention to the things voters pay attention to and the actions they suggest in the survey, political Parties can hope to steadily increase voter satisfaction, aiding them in the reelection process.

With further research on this study and the other ces datasets, these conclusions can help policymakers determine, definitively, which policies they can enact to result in the greatest level of voter satisfaction. 


# Limitations

The most apparent limitation of the dataset is the fact that it is a survey where not all answers can be checked for truthfulness. That is, we must rely on the people taking this survey to provide honest answers that we can depend on for our analysis. With questions like 'who did you vote for?' or 'who will you vote for?' it would be very easy for someone taking this survey to either lie or alter their decision which would affect the true outcome of this study. Since, however, there is no realistic way of checking for this type of thing, we must operate under the assumption that we are being provided with completely true data.
Aside from falsified answers to the survey, questions like 'how much attention do you pay to the environment' can be quite vague and possibly result in a number of people answering the question in different ways leading to inconsistent answers. Furthermore, the number of questions that were unanswered by voters also makes it extremely difficult to get a full picture of what voters of each political Party most want.

Further data cleaning could have been conducted such as removing responses like 'Seen but not answered' to help to make the displays of data appear more full. Additionally, I would have liked to avoid the overlapping of bar names for my graphs as I felt it took a great deal away from the study by leaving it to look extremely cluttered.
Additional factors I could have selected from the ces2015_web could have helped provide more insight into what is most important to voters and what policymakers should focus on. It would have been useful to cover more questions due to the lack of important information gained regarding what policies the Conservative Party should focus on. For example, choosing another topic such as views on immigration or crime rates may have allowed for me to make another suggestion as to how the Conservative Party can gain voter satisfaction. 
It would have been very nice to be able to study the ces2019_web data due to its more recent results, however, due to some error I must have made while loading and cleaning the data for it I became unable to access it. Regardless, I feel that the ces2015_web data should still provide a clear indication of what voters consider important and what they feel should be done about it.

  
# Appendix

Code for this study can be found at: https://github.com/jordanregis/TJRProjCode/blob/master/TJRProjCode.Rmd

# References

We used:

@citeR

R Core Team (2020). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna,
  Austria. URL https://www.R-project.org/.

@citetidyverse

 Wickham et al., (2019). Welcome to the tidyverse. Journal of Open
  Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

@citeskimr

  Elin Waring, Michael Quinn, Amelia McNamara, Eduardo Arino de la
  Rubia, Hao Zhu and Shannon Ellis (2020). skimr: Compact and
  Flexible Summaries of Data. https://docs.ropensci.org/skimr
  (website), https://github.com/ropensci/skimr.

@citelabelled

  Joseph Larmarange (2020). labelled: Manipulating Labelled Data. R
  package version 2.7.0. http://larmarange.github.io/labelled/

@citecesR

  Paul A. Hodgetts and Rohan Alexander (2020). cesR: Access the CES
  Datasets a Little Easier.. R package version 0.1.0.

@citevisdat

Tierney N (2017). “visdat: Visualising Whole Data Frames.” _JOSS_,
*2*(16), 355. doi: 10.21105/joss.00355 (URL:
https://doi.org/10.21105/joss.00355), <URL:
http://dx.doi.org/10.21105/joss.00355>.

@citeggplot2

  H. Wickham. ggplot2: Elegant Graphics for Data Analysis.
  Springer-Verlag New York, 2016.

@citedevtools

  Hadley Wickham, Jim Hester and Winston Chang (2020).
  devtools: Tools to Make Developing R Packages Easier.
  https://devtools.r-lib.org/,
  https://github.com/r-lib/devtools.

(Note: It appears that below this line, the references have repeated themselves for a reason I could not work out)